### 2019-09-03 docs/detailed-example.Rmd
Tidy up

---
title: "AIMS Data Platform R Client Notebook Example"
output: html_notebook
---

[Back](index)

This example assumes you have obtained an *AIMS Data Platform API Key*.  See [here](index#usage-of-aims-dataplatform-api-key) to see what that means.  To download and run this example interactively in *R Studio* click on `Code -> Download Rmd` in the top right corner of this document.

Once you have obtained your API Key past it in the following code chunk and it will be used throughout this tutorial:

```{r}
# paste your key where it says api-key-for-r-notebook
MY_API_KEY <- "api-key-for-r-notebook"

```

Install R Client
----------------

*AIMS Data Platform R Client* can be installed directly from [github](https://github.com/aims/data-platform-r) using the following code:

```{r}
install.packages("devtools")
devtools::install_git("https://github.com/AIMS/data-platform-r")
library(aimsdataplatform)

```

Discover a dataset
------------------

Suppose we want to query a data series and plot it then the procedure might be:
1. Examine documentation and establish query filters
2. Perform data download using `getData`
3. Use R `plot` to create a chart

We decide to query the [AIMS Weather](https://aims.github.io/data-platform/weather) data set based on a data series, then we are guaranteed of getting information from one site for one metric.

To discover a data set we look up the filter values available for a series:

```{r}
getFilterValues('10.25845/5c09bf93f315d', 'series')
```

The benefits to choosing a data series is that it comes from one location and parameter type, making the data easy to plot.  If we did not choose a data series from the *AIMS Weather* data set we might have to specify additional parameters to ensure the data is as expected.

Finally our filters might look like the following:

Variable  | Value                  | Description
----------|------------------------|-------------------------------------------------------
dio       | 10.25845/5c09bf93f315d | DOI for AIMS Weather found [here](https://aims.github.io/data-platform)
series    | 64                     | Found [here](https://aims.github.io/data-platform/weather/series), Davies Reef Air Temperature data series
from-date | '2018-01-01'           | We want to start charting on 1/1/2018
thru-date | '2018-01-10'           | We are plotting 10 days of data

Query and Plot Dataset
----------------------

After deciding on query parameters we plug the series id into a `getData` function:

```{r}
results <- getData('10.25845/5c09bf93f315d', apiKey=MY_API_KEY, filters=list('series'=64, 'from-date'='2018-01-01', 'thru-date'='2018-01-10'))

```

Now the results have been stored in the `results` we can attempt to plot.

```{r}
plot(results$dataFrame$time, results$dataFrame$qc_value)

```

Ok we have a plot, fairly unpleasant looking, that should be a line chart, also we should fix the time labels by putting the X values into an axis command:

```{r}
plot(results$dataFrame$time, results$dataFrame$qc_value, type='l')

```

Lets add some proper labels:

```{r}
plot(results$dataFrame$time, results$dataFrame$qc_value, type='l', xlab="Time", ylab="Air Temperature")

```

Finally lets make the colour green and add a main title.

```{r}
plot(results$dataFrame$time, results$dataFrame$qc_value, type='l', xlab="Time", ylab="Air Temperature", main="Davies Reef Platform Air Temperature in 2018", col='green')

```

And add another series:

```{r}
results2 <- getData('10.25845/5c09bf93f315d', apiKey=MY_API_KEY, filters=list('series'=87929, 'from-date'='2018-01-01', 'thru-date'='2018-01-10'))
plot(results$dataFrame$time, results$dataFrame$qc_value, type='l', xlab="January 2018", ylab="Air Temperature", main="Air Temperature at AIMS Weather Stations in 2018", col='green')
lines(results2$dataFrame$time, results2$dataFrame$qc_value, col="blue")

```

Now a legend to distinguish data sets

```{r}
plot(results$dataFrame$time, results$dataFrame$qc_value, type='l', xlab="Time", ylab="Air Temperature", main="Air Temperature at AIMS Weather Stations in 2018", col='green')
lines(results2$dataFrame$time, results2$dataFrame$qc_value, col="blue")
legend("bottomright", c("Davies Reef", "Bramble Cay"), fill=c("green", "blue"))

```

Finally citations can be added to our document with the following code:

```{r}
results$citation
results2$citation

```

Bibliography
------------

1. `r results$citation`

2. `r results2$citation`