### 2021-02-11 paper/paper.Rmd
added paper skeleton

---
title: 'dataaimsr: A Client for the Australian Institute of Marine Science Data Platform API which provides easy access to AIMS Data Platform'
tags:
  - R
  - sea surface temperature
  - weather stations
  - long-term environmental monitoring
  - Australia
  - API
authors:
  - name: Diego R. Barneche
    orcid: 0000-0002-4568-2362
    affiliation: "1, 2"
  - name: Greg Coleman
    affiliation: "3"
  - name: Jeffrey L. Sheehan
    affiliation: "3"
  - name: Duncan Fermor
    affiliation: "3"
  - name: Murray Logan
    affiliation: "3"
  - name: Mark Rehbein
    affiliation: "3"
affiliations:
 - name: Australian Institute of Marine Science, Crawley, WA 6009, Australia
   index: 1
 - names: The Indian Ocean Marine Research Centre, The University of Western Australia, Crawley, WA 6009, Australia
   index: 2
 - name: Australian Institute of Marine Science, Townsville, Qld 4810, Australia
   index: 3

citation_author: Barneche et al.
date: "`r Sys.Date()`"
bibliography: paper.bib
output:
  my_modified_joss:
    fig_caption: yes
csl: apa.csl
journal: JOSS
---

# Summary

# Background and Statement of need

# Technical details and Usage

## Temperature-logger dataset

### Summary methods

### Datasets

## Weather station datasets

### Summary methods

### Datasets

# Future directions

# Acknowledgements

The development of `dataimsr` was supported by ... Names to be added to the list.

# References