### 2023-05-19 test_running_updating.Rmd
updates to tests

# How to run or recreate tests

From `dataaimsr/`, run

```{r}
library(testthat)
library(roxygen2)
library(devtools)
system("rm -rf tests/api")
roxygen2::roxygenise()
devtools::install(build_vignettes = FALSE)
testthat::test_local(".")
all_mock_dirs <- dir("tests/api", full.names = TRUE, recursive = TRUE) |>
  dirname() |>
  unique()
new_mock_dirs <- gsub(
  "api.aims.gov.au/data-v2.0/10.25845", "a", all_mock_dirs, fixed = TRUE
)
for (i in seq_along(all_mock_dirs)) {
  system(paste0("mkdir -p ", new_mock_dirs[i]))
  system(paste0(
    "cp -R ", all_mock_dirs[i], "/* ", new_mock_dirs[i], "/" , collapse = ""
  ))
  to_rm <- dirname(all_mock_dirs[i]) |>
    dirname() |>
    dirname()
  system(paste0("rm -rf ", to_rm))
}
```

This will delete all existing mock files and regenerate them. Then return to root directory and check if it all passes:

```{r}
rm(list = ls())
rcmdcheck::rcmdcheck(
  args = c("--no-manual", "--as-cran"), build_args = "--resave-data=best",
  error_on = "warning", check_dir = "check"
)
```