### 2023-05-31 test_running_updating.Rmd
updated tests following API update

# How to run or recreate tests

From `dataaimsr/`, run

```{r}
library(httptest)
library(testthat)
library(roxygen2)
library(devtools)
system("rm -rf tests/api")
roxygen2::roxygenise()
devtools::install(build_vignettes = FALSE)
testthat::test_local(".")
```

The above block will delete all existing mock files and regenerate them. Then return to root directory and check if it all passes:

```{r}
rm(list = ls())
rcmdcheck::rcmdcheck(
  args = c("--no-manual", "--as-cran"), build_args = "--resave-data=best",
  error_on = "warning", check_dir = "check"
)
```